Massively Parallel Monte Carlo Methods for Discrete Linear and Nonlinear Systems

### 2012-11-05 doc/prelim/presentation/prelim_presentation.tex
adding spacing to slides

  \begin{itemize}
  \item Predictive modeling and simulation enhances engineering
    capability
  \item Modern work focused on this task leverages multiple physics
    simulation (CASL, NEAMS)
  \item New hardware drives algorithm development (petascale and
    exascale)
  \item Monte Carlo methods have the potential to provide great
    improvements that permit finer simulations and better mapping to
    future hardware
  \item A set of massively parallel Monte Carlo methods is proposed to
    advance multiple physics simulation on contemporary and future
    leadership class machines
  \end{itemize}

@@

  \begin{itemize}
  \item Modern hardware is moving in two directions (Kogge,2011):
    \begin{itemize}
    \item Lightweight machines
    \item Heterogeneous machines
    \item Both characterized by low power and high concurrency
    \end{itemize}
  \item Some issues:
    \begin{itemize}
    \item Higher potential for both soft and hard failures (DOE,2012)
    \item Memory restrictions are expected with a continued decrease
      in memory/FLOPS
    \end{itemize}
  \item Potential resolution from Monte Carlo:
    \begin{itemize}
    \item Soft failures buried within the tally variance
    \item Hard failures are high variance events
    \item Memory savings over conventional methods
    \end{itemize}
  \end{itemize}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Research Outline}
  \begin{itemize}
  \item Parallelization of Monte Carlo methods for discrete systems
    \begin{itemize}
    \item Parallel strategies taken from modern reactor physics
      methods
    \item Research is required to explore varying parallel strategies
    \item Scalability is of concern
    \end{itemize}
  \item Development of a nonlinear solver for discrete systems
    leveraging Monte Carlo
    \begin{itemize}
    \item Application to nonlinear problems of interest
    \item Memory benefits
    \item Performance benefits
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {A Requirement} Assert that $\ve{A}$ is \textit{nonsingular}. The
    solution is then:
    \[
    \ve{x} = \ve{A}^{-1}\ve{b}
    \]
  \end{beamerboxesrounded}

@@

  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Search Subspace $\mathcal{K}$} Extract the solution from the
    search subspace:
    \[
    \tilde{\ve{x}} = \ve{x}_0 +
    \boldsymbol{\delta},\ \boldsymbol{\delta} \in \mathcal{K}
    \]
  \end{beamerboxesrounded}

  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Constraint Subspace $\mathcal{L}$} Constrain the extraction with
    the constraint subspace by asserting orthogonality with the
    residual:
    \[
    \langle \tilde{\ve{r}},\ve{w} \rangle = 0,\ \forall \ve{w} \in
    \mathcal{L}
    \]  
  \end{beamerboxesrounded}

@@

  \begin{itemize}
  \item For GMRES (Saad, 1986):
  \end{itemize}
  \[
  \mathcal{L} = \ve{A} \mathcal{K}_m(\ve{A},\ve{r}_0)
  \]

  \begin{itemize}
  \item Yields the normal system $\ve{A}^T\ve{A}\ve{x} =
    \ve{A}^T\ve{b}$
  \item Must generate an orthonormal basis for
    $\mathcal{K}_m(\ve{A},\ve{r}_0)$
  \item Typically choose a Gram-Schmidt-like procedure such as Arnoldi
    or Lanzcos
  \item Short and long recurrence relations available for
    orthogonalization
  \end{itemize}

@@

  \begin{itemize}
  \item Parallel dot product
  \end{itemize}
  \[
  d_l = \ve{y}_l \cdot \ve{x}_l,\ d_g = \sum_p d_l
  \]

  \begin{itemize}
  \item Parallel vector norm
  \end{itemize}
  \[
  ||x||_{\infty,l} = \max_n \ve{y}[n],\ \forall n \in [1,N_l]
  \]
  \[
  ||x||_{\infty,g} = \max_p ||x||_{\infty,l}
  \]

@@

  \begin{itemize}
  \item Widely used in practice
  \item Krylov methods require only the action of the operator
  \item Parallelism achieved through a handful of operations
  \item Global reduction operations observed not to impede scalability
    (Gropp,2001)
    \begin{itemize}
    \item Dot product
    \item Vector norms
    \end{itemize}
  \item Nearest neighbor computations have poor algorithmic strong
    scaling
    \begin{itemize}
    \item Matrix-vector multiply
    \item Weak scaling is better
    \end{itemize}
  \item Short and long recurrence relations available for
    orthogonalization
  \end{itemize}

@@

  \begin{itemize}
  \item First proposed by J. Von Neumann and S.M. Ulam in the 1940's
  \item Earliest published reference in 1950
  \item General lack of published work
  \item Modern work by Evans and others has yielded new applications
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item Generate the \textit{Neumann series}
  \end{itemize}

@@

  \begin{itemize}
  \item Require $\rho(\ve{H}) < 1$ for convergence
  \end{itemize}

@@

  \begin{itemize}
  \item Define a sequence of state transitions
  \end{itemize}

@@

  \begin{itemize}
  \item Define the \textit{Neumann-Ulam decomposition}\footnote{The
    Hadamard product $\ve{A} = \ve{B} \circ \ve{C}$ is defined
    element-wise as $a_{ij} = b_{ij} c_{ij}$.}
  \end{itemize}

@@

  \begin{itemize}
  \item Generate an expectation value for the solution
  \end{itemize}

@@

  \begin{itemize}
  \item Set the adjoint constraint
  \end{itemize}

@@

  \[
  \ve{y} = (\ve{I} - \ve{H}^T)^{-1} \ve{d}
  \]

  \[
  \ve{y} = \sum_{k=0}^{\infty} (\ve{H}^T)^k\ve{d}
  \]

  \begin{itemize}
  \item Expand the series
  \end{itemize}

@@

  \begin{itemize}
  \item Pick another constraint to yield the original solution
  \end{itemize}

@@

  \begin{itemize}
  \item Build the estimator and expectation value
  \end{itemize}

@@

  \begin{itemize}
  \item Neumann-Ulam methods bound by the Central Limit Theorem
  \item Build on Halton's 1962 Sequential Monte Carlo method
  \item Adjoint Neumann-Ulam solver computes the correction
  \item Decouples MC error from solution error, exponential convergence
  \item Demonstrated by Evans to be competitive with Krylov methods
  \end{itemize}

@@

  \begin{itemize}
  \item Yields a preconditioned MCSA iteration with no in-state
    transitions
  \end{itemize}

@@

  \begin{itemize}
  \item Published work to date has used a physics-based formulation
    \begin{itemize}
    \item Radiation transport equations used as model system
    \item Transition probabilities built from problem-specific
      parameters
    \item Probabilities and weights must be re-derived for each new
      equation set
    \end{itemize}
  \item Desire a generalization for all linear operator equations
    \begin{itemize}
    \item Requires a general parallel framework
    \item Requires implementation with a general linear algebra
      framework
    \item Operator, vector, and graph abstractions
    \end{itemize}
  \item Neumann-Ulam solvers and MCSA implemented using the Trilinos
    Petra frameworks
  \item Can be leveraged in modern physics implementations
  \end{itemize}

@@

  \begin{itemize}
  \item No literature observed for parallel Neumann-Ulam solvers
  \item Numerous references for modern parallel Monte Carlo methods in
    reactor physics
  \item Build a strategy for applying modern methods to the
    Neumann-Ulam method
  \item MCSA iteration-level parallelism comes from parallel
    matrix/vector operations
  \end{itemize}

@@

  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Brunner's Work (2006 and 2009)}

    \begin{itemize}
    \item Looked at 4 communication patterns:
      \begin{itemize}
      \item Fully-locking synchronous
      \item asynchronous-send/synchronous-receive
      \item master/slave
      \item binary tree master/slave
      \end{itemize}
    \item Binary tree master/slave performed best but race conditions observed
    \item A later improvement to their work showed a fully asynchronous
      pattern to work best
    \item Poor scaling for unbalanced problems for all schemes
    \end{itemize}

@@

  \begin{itemize}
  \item The amount of domain leakage dictates communication
  \item Per Siegel's 2012 work, define a leakage fraction
  \end{itemize}

@@

  \begin{itemize}
  \item Siegel observed $\lambda$ to be bound empirically to the mean
    free path of the system
  \item Optimum ratio of local to global domain size exists
  \item Experiments show feasibility for large-scale calculations
    \begin{itemize}
    \item Not bandwidth or latency limited in load-balanced case
    \end{itemize}
    \pause
  \item How do we define $\lambda$ empirically for linear operator
    equations? 
  \end{itemize}

@@

Questions
  \begin{itemize}
  \item How much domain overlap is suitable for linear operator
    equations? 
  \item Is memory a limitation for overlap and replication?
  \item Do the linear operator properties help select overlap?
  \item Will full-clip roulette perturb the MCSA solution?
  \item How much does MSOD facilitate scaling?
  \end{itemize}

@@

  \begin{itemize}
  \item Richardson iteration and residual computation require parallel
    matrix-vector multiply and parallel vector update
  \item This work will generate a parallel adjoint Neumann-Ulam solver
  \item Application of correction requires parallel vector update
  \item Convergence checks through parallel norm computation
  \end{itemize}

@@

  \begin{itemize}
  \item Many multiphysics problems of interest are nonlinear
  \item Segregated methods lack the consistency of fully implicit
    methods
  \item Newton methods often leverage Krylov solvers
    \begin{itemize}
    \item Robust implementations
    \item No operator required
    \item Memory intensive
    \end{itemize}
  \item Monte Carlo methods need the full operator
  \item Automatic construction of the linear operator available
    \begin{itemize}
    \item Ideal for Monte Carlo
    \item Relaxes memory requirements
    \item Potential scaling improvements
    \item Resiliency benefits
    \end{itemize}
  \end{itemize}

@@

  \begin{itemize}
  \item Choose a Krylov method to solve for the Newton correction
  \item GMRES with a long recurrence relation observed as more
    robust (Knoll,2004)
  \item Generates a monotonically decreasing residual from
    maintaining the optimization and orthogonality conditions
  \end{itemize}

  \pause Where does the Jacobian come from?
  \begin{itemize}
  \item The Jacobian can come from hand-coded derivatives
    \begin{itemize}
    \item Tedious and error prone
    \item Repeated for each equation set and hard to do for
      multiphysics
    \end{itemize}
  \end{itemize}

@@

  \begin{itemize}
  \item Methods verification
  \item Numerical experiments
  \item Challenge problem
  \end{itemize}

@@

  \begin{itemize}
  \item Analytic solution to the heat equation for the linear methods
  \item Sequence of Navier-Stokes benchmarks for the nonlinear methods
    \begin{itemize}
    \item Thermal convection cavity problem (De Vahl Davis, 1983)
    \item Lid driven cavity problem (Ghia et al., 1982)
    \item Backward-Facing step problem (Gartling, 1990)
    \end{itemize}
  \item Tuning benchmark parameters varies the strength of
    nonlinearities
  \end{itemize}

  \begin{columns}
    \begin{column}{0.5\textwidth}
      \[
      \rho \ve{u} \cdot \nabla \ve{u} - \nabla \cdot \ve{T} - \rho
      \ve{g} = \ve{0}
      \]
      \[
      \nabla \cdot \ve{u} = 0
      \]
      \[
      \rho C_p \ve{u} \cdot \nabla T + \nabla \cdot \ve{q} = 0
      \]
    \end{column}

@@

  \pause
  \textbf{Domain Replication Studies for Parallel Neumann-Ulam Method}
  \begin{itemize}
  \item Feasibility from a memory perspective
  \item Impact on scalability
  \item Impact on load balancing
  \end{itemize}

@@

  \pause
  \textbf{Parallel Performance and Numerical Accuracy Studies for FANM
    Method}
  \begin{itemize}
  \item Feasibility for problems of interest
  \item Memory usage vs. Newton-Krylov methods
  \item Scalability vs. Newton-Krylov methods
  \end{itemize}

@@

  \begin{itemize}
  \item Problems of interest are coupled large scale problems in
    reactor physics
  \item The Consortium for Advanced Simulation of LWRs (CASL) is a
    modeling and simulation program aimed at:
    \begin{itemize}
      \item Higher power uprates and efficiency
      \item Higher burn-up
      \item Predictive accident scenario analysis
    \end{itemize}
  \item CASL and industry partners identified challenge problems:
    \begin{itemize}
    \item Departure from nucleate boiling
    \item Grid-to-rod-fretting
    \end{itemize}
  \item Our challenge problem should reflect how this work aids the
    solution of these problems
  \end{itemize}

@@

  \begin{itemize}
  \item Proposed research and development of new Monte Carlo methods
    \begin{itemize}
    \item Parallelization of Monte Carlo methods for linear systems
    \item FANM for nonlinear systems
    \end{itemize}
  \item Verification through benchmarks
  \item Numerical experiments for understanding
  \item Challenge problem for application
  \item Directed towards hard problems in nuclear reactor analysis:
    \begin{itemize}
    \item Application to multiphysics
    \item Potential improvements for scalability
    \item Potential improvements for memory consumption
    \item Looking forward to exascale
    \end{itemize}
  \end{itemize}

@@

  Thanks to...\footnote{This work was performed under
        appointment to the Nuclear Regulatory Commission Fellowship
        program at the University of Wisconsin - Madison Department of
        Engineering Physics.}

@@

>>>>

  \begin{itemize}
  \item Predictive modeling and simulation enhances engineering
    capability
    \medskip
  \item Modern work focused on this task leverages multiple physics
    simulation (CASL, NEAMS)
    \medskip
  \item New hardware drives algorithm development (petascale and
    exascale)
    \medskip
  \item Monte Carlo methods have the potential to provide great
    improvements that permit finer simulations and better mapping to
    future hardware
    \medskip
  \item A set of massively parallel Monte Carlo methods is proposed to
    advance multiple physics simulation on contemporary and future
    leadership class machines
  \end{itemize}

@@

  \begin{itemize}
  \item Modern hardware is moving in two directions (Kogge,2011):
    \begin{itemize}
    \item Lightweight machines
    \item Heterogeneous machines
    \item Both characterized by low power and high concurrency
    \end{itemize}
    \medskip \medskip
  \item Some issues:
    \begin{itemize}
    \item Higher potential for both soft and hard failures (DOE,2012)
    \item Memory restrictions are expected with a continued decrease
      in memory/FLOPS
    \end{itemize}
    \medskip \medskip
  \item Potential resolution from Monte Carlo:
    \begin{itemize}
    \item Soft failures buried within the tally variance
    \item Hard failures are high variance events
    \item Memory savings over conventional methods
    \end{itemize}
  \end{itemize}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Research Outline}
  \begin{itemize}
  \item Parallelization of Monte Carlo methods for discrete systems
    \begin{itemize}
    \item Parallel strategies taken from modern reactor physics
      methods
    \item Research is required to explore varying parallel strategies
    \item Scalability is of concern
    \end{itemize}
    \medskip \medskip \medskip
  \item Development of a nonlinear solver for discrete systems
    leveraging Monte Carlo
    \begin{itemize}
    \item Application to nonlinear problems of interest
    \item Memory benefits
    \item Performance benefits
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \medskip \medskip 
  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {A Requirement} Assert that $\ve{A}$ is \textit{nonsingular}. The
    solution is then:
    \[
    \ve{x} = \ve{A}^{-1}\ve{b}
    \]
  \end{beamerboxesrounded}

@@

  \medskip \medskip
  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Search Subspace $\mathcal{K}$} Extract the solution from the
    search subspace:
    \[
    \tilde{\ve{x}} = \ve{x}_0 +
    \boldsymbol{\delta},\ \boldsymbol{\delta} \in \mathcal{K}
    \]
  \end{beamerboxesrounded}

  \medskip \medskip
  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Constraint Subspace $\mathcal{L}$} Constrain the extraction with
    the constraint subspace by asserting orthogonality with the
    residual:
    \[
    \langle \tilde{\ve{r}},\ve{w} \rangle = 0,\ \forall \ve{w} \in
    \mathcal{L}
    \]  
  \end{beamerboxesrounded}

@@

  \begin{itemize}
  \item For GMRES (Saad, 1986): $\mathcal{L} = \ve{A}
    \mathcal{K}_m(\ve{A},\ve{r}_0)$
  \end{itemize}

  \medskip \medskip \medskip
  \small{
  \begin{itemize}
  \item Yields the normal system $\ve{A}^T\ve{A}\ve{x} =
    \ve{A}^T\ve{b}$
    \medskip
  \item Must generate an orthonormal basis for
    $\mathcal{K}_m(\ve{A},\ve{r}_0)$
    \medskip
  \item Typically choose a Gram-Schmidt-like procedure such as Arnoldi
    or Lanzcos
    \medskip
  \item Short and long recurrence relations available for
    orthogonalization
  \end{itemize}
  }

@@

  \medskip \medskip
  \begin{itemize}
  \item Parallel dot product
  \end{itemize}
  \[
  d_l = \ve{y}_l \cdot \ve{x}_l,\ d_g = \sum_p d_l
  \]

  \medskip \medskip
  \begin{itemize}
  \item Parallel vector norm
  \end{itemize}
  \[
  ||x||_{\infty,l} = \max_n \ve{y}[n],\ \forall n \in [1,N_l]
  \]
  \[
  ||x||_{\infty,g} = \max_p ||x||_{\infty,l}
  \]

@@

  \begin{itemize}
  \item Widely used in practice
    \medskip
  \item Krylov methods require only the action of the operator
    \medskip
  \item Parallelism achieved through a handful of operations
    \medskip
  \item Global reduction operations observed not to impede scalability
    (Gropp,2001)
    \begin{itemize}
    \item Dot product
    \item Vector norms
    \end{itemize}
    \medskip
  \item Nearest neighbor computations have poor algorithmic strong
    scaling
    \begin{itemize}
    \item Matrix-vector multiply
    \item Weak scaling is better
    \end{itemize}
    \medskip
  \item Short and long recurrence relations available for
    orthogonalization
  \end{itemize}

@@

  \begin{itemize}
  \item First proposed by J. Von Neumann and S.M. Ulam in the 1940's
    \medskip \medskip
  \item Earliest published reference in 1950
    \medskip \medskip
  \item General lack of published work
    \medskip \medskip
  \item Modern work by Evans and others has yielded new applications
  \end{itemize}
\end{frame}

@@

  \medskip
  \begin{itemize}
  \item Generate the \textit{Neumann series}
  \end{itemize}

@@

  \medskip
  \begin{itemize}
  \item Require $\rho(\ve{H}) < 1$ for convergence
  \end{itemize}

@@

  \medskip
  \begin{itemize}
  \item Define a sequence of state transitions
  \end{itemize}

@@

  \medskip
  \begin{itemize}
  \item Define the \textit{Neumann-Ulam decomposition}\footnote{The
    Hadamard product $\ve{A} = \ve{B} \circ \ve{C}$ is defined
    element-wise as $a_{ij} = b_{ij} c_{ij}$.}
  \end{itemize}

@@

  \medskip \medskip
  \begin{itemize}
  \item Generate an expectation value for the solution
  \end{itemize}

@@

  \medskip
  \begin{itemize}
  \item Set the adjoint constraint
  \end{itemize}

@@

  \[
  \ve{y} = (\ve{I} - \ve{H}^T)^{-1} \ve{d} = \sum_{k=0}^{\infty}
  (\ve{H}^T)^k\ve{d}
  \]

  \medskip
  \begin{itemize}
  \item Expand the series
  \end{itemize}

@@

  \medskip
  \begin{itemize}
  \item Pick another constraint to yield the original solution
  \end{itemize}

@@

  \medskip
  \begin{itemize}
  \item Build the estimator and expectation value
  \end{itemize}

@@

  \medskip \medskip
  \begin{itemize}
  \item Neumann-Ulam methods bound by the Central Limit Theorem
  \item Build on Halton's 1962 Sequential Monte Carlo method
  \item Adjoint Neumann-Ulam solver computes the correction
  \item Decouples MC error from solution error, exponential convergence
  \item Demonstrated by Evans to be competitive with Krylov methods
  \end{itemize}

@@

  \medskip
  \begin{itemize}
  \item Yields a preconditioned MCSA iteration with no in-state
    transitions
  \end{itemize}

@@

  \begin{itemize}
  \item Published work to date has used a physics-based formulation
    \begin{itemize}
    \item Radiation transport equations used as model system
    \item Transition probabilities built from problem-specific
      parameters
    \item Probabilities and weights must be re-derived for each new
      equation set
    \end{itemize}
    \medskip \medskip
  \item Desire a generalization for all linear operator equations
    \begin{itemize}
    \item Requires a general parallel framework
    \item Requires implementation with a general linear algebra
      framework
    \item Operator, vector, and graph abstractions
    \end{itemize}
    \medskip \medskip
  \item Neumann-Ulam solvers and MCSA implemented using the Trilinos
    Petra frameworks
    \medskip \medskip
  \item Can be leveraged in modern physics implementations
  \end{itemize}

@@

  \begin{itemize}
  \item No literature observed for parallel Neumann-Ulam solvers
    \medskip \medskip
  \item Numerous references for modern parallel Monte Carlo methods in
    reactor physics
    \medskip \medskip
  \item Build a strategy for applying modern methods to the
    Neumann-Ulam method
    \medskip \medskip
  \item MCSA iteration-level parallelism comes from parallel
    matrix/vector operations
  \end{itemize}

@@

  \medskip \medskip
  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Brunner's Work (2006 and 2009)}

    \begin{itemize}
    \item Looked at 4 communication patterns:
      \begin{itemize}
      \item Fully-locking synchronous
      \item Asynchronous-send/synchronous-receive
      \item Master/slave
      \item Binary tree master/slave
      \end{itemize}
    \item Binary tree master/slave performed best but race conditions observed
    \item A later improvement to their work showed a fully asynchronous
      pattern to work best
    \item Poor scaling for unbalanced problems for all schemes
    \end{itemize}

@@

  \begin{itemize}
  \item The amount of domain leakage dictates communication
    \medskip
  \item Per Siegel's 2012 work, define a leakage fraction
  \end{itemize}

@@

  \medskip
  \begin{itemize}
  \item Siegel observed $\lambda$ to be bound empirically to the mean
    free path of the system
    \medskip
  \item Optimum ratio of local to global domain size exists
    \medskip
  \item Experiments show feasibility for large-scale calculations
    \begin{itemize}
    \item Not bandwidth or latency limited in load-balanced case
    \end{itemize}
    \pause
    \medskip
  \item How do we define $\lambda$ empirically for linear operator
    equations? 
  \end{itemize}

@@

  \medskip \medskip \medskip
Questions
  \begin{itemize}
  \item How much domain overlap is suitable for linear operator
    equations? 
  \item Is memory a limitation for overlap and replication?
  \item Do the linear operator properties help select overlap?
  \item Will full-clip roulette perturb the MCSA solution?
  \item How much does MSOD facilitate scaling?
  \end{itemize}

@@

  \medskip \medskip
  \begin{itemize}
  \item Richardson iteration and residual computation require parallel
    matrix-vector multiply and parallel vector update
  \item This work will generate a parallel adjoint Neumann-Ulam solver
  \item Application of correction requires parallel vector update
  \item Convergence checks through parallel norm computation
  \end{itemize}

@@

  \begin{itemize}
  \item Many multiphysics problems of interest are nonlinear
    \medskip
  \item Segregated methods lack the consistency of fully implicit
    methods
    \medskip
  \item Newton methods often leverage Krylov solvers
    \begin{itemize}
    \item Robust implementations
    \item No operator required
    \item Memory intensive
    \end{itemize}
    \medskip
  \item Monte Carlo methods need the full operator
    \medskip
  \item Automatic construction of the linear operator available
    \begin{itemize}
    \item Ideal for Monte Carlo
    \item Relaxes memory requirements
    \item Potential scaling improvements
    \item Resiliency benefits
    \end{itemize}
  \end{itemize}

@@

  \begin{itemize}
  \item Choose a Krylov method to solve for the Newton correction
    \medskip
  \item GMRES with a long recurrence relation observed as more
    robust (Knoll,2004)
    \medskip
  \item Generates a monotonically decreasing residual from
    maintaining the optimization and orthogonality conditions
  \end{itemize}

  \medskip \medskip \medskip
  \pause Where does the Jacobian come from?
  \begin{itemize}
  \item The Jacobian can come from hand-coded derivatives
    \begin{itemize}
    \item Tedious and error prone
    \item Repeated for each equation set and hard to do for
      multiphysics
    \end{itemize}
  \end{itemize}

@@

  \large{
  \begin{itemize}
  \item Methods verification
  \medskip \medskip \medskip \medskip \medskip
  \item Numerical experiments
    \medskip \medskip \medskip \medskip \medskip
  \item Challenge problem
  \end{itemize}
  }

@@

  \begin{itemize}
  \item Analytic solution to the heat equation for the linear methods
    \medskip
  \item Sequence of Navier-Stokes benchmarks for the nonlinear methods
    \begin{itemize}
    \item Thermal convection cavity problem (De Vahl Davis, 1983)
    \item Lid driven cavity problem (Ghia et al., 1982)
    \item Backward-Facing step problem (Gartling, 1990)
    \end{itemize}
    \medskip
  \item Tuning benchmark parameters varies the strength of
    nonlinearities
  \end{itemize}

  \medskip  \medskip
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \[
      \rho \ve{u} \cdot \nabla \ve{u} - \nabla \cdot \ve{T} - \rho
      \ve{g} = \ve{0}
      \]
      \[
      \nabla \cdot \ve{u} = 0
      \]
      \[
      \rho C_p \ve{u} \cdot \nabla T + \nabla \cdot \ve{q} = 0
      \]
    \end{column}

@@

  \medskip  \medskip
  \pause
  \textbf{Domain Replication Studies for Parallel Neumann-Ulam Method}
  \begin{itemize}
  \item Feasibility from a memory perspective
  \item Impact on scalability
  \item Impact on load balancing
  \end{itemize}

@@

  \medskip  \medskip
  \pause
  \textbf{Parallel Performance and Numerical Accuracy Studies for FANM
    Method}
  \begin{itemize}
  \item Feasibility for problems of interest
  \item Memory usage vs. Newton-Krylov methods
  \item Scalability vs. Newton-Krylov methods
  \end{itemize}

@@

  \begin{itemize}
  \item Problems of interest are coupled large scale problems in
    reactor physics
    \medskip
  \item The Consortium for Advanced Simulation of LWRs (CASL) is a
    modeling and simulation program aimed at:
    \begin{itemize}
      \item Higher power uprates and efficiency
      \item Higher burn-up
      \item Predictive accident scenario analysis
    \end{itemize}
    \medskip
  \item CASL and industry partners identified challenge problems:
    \begin{itemize}
    \item Departure from nucleate boiling
    \item Grid-to-rod-fretting
    \end{itemize}
    \medskip
  \item Our challenge problem should reflect how this work aids the
    solution of these problems
  \end{itemize}

@@

  \begin{itemize}
  \item Proposed research and development of new Monte Carlo methods
    \begin{itemize}
    \item Parallelization of Monte Carlo methods for linear systems
    \item FANM for nonlinear systems
    \end{itemize}
    \medskip
  \item Verification through benchmarks
    \medskip
  \item Numerical experiments for understanding
    \medskip
  \item Challenge problem for application
    \medskip
  \item Directed towards hard problems in nuclear reactor analysis:
    \begin{itemize}
    \item Application to multiphysics
    \item Potential improvements for scalability
    \item Potential improvements for memory consumption
    \item Looking forward to exascale
    \end{itemize}
  \end{itemize}

@@

  Thanks to...

@@

  \medskip \medskip \medskip \medskip \medskip \medskip \medskip

  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Support} 
    This work was performed under appointment to the Nuclear Regulatory
    Commission Fellowship program at the University of Wisconsin - Madison
    Department of Engineering Physics.
  \end{beamerboxesrounded}
